$p$-adic precision

### 2014-02-06 padicprec.tex
Typos corrected and one remark added

We now address the case of a multiple roots. Let $P_0$ be a monic 
polynomial of degree $d$ and $\alpha_0 \in K$ be a root of $P_0$ having 
multiplicity $m > 1$. Due to the multiplicity, it is no longer possible 
to follow to root $\alpha_0$ in a neighborhood of $P_0$. But it is if we 
restrict ourselves to polynomials which have a root of multiplicity $m$ 
close to $\alpha_0$. More precisely, let us consider the locus $\mathcal 
V_m$ consisting of monic polynomials of degree $d$ having a root of 
multiplicity at least $m$. It is a Zariski close subset of $K_d[X]$ and 
it contains by assumption the polynomial $P_0$ we started with. 
Moreover, the irreducible components of $\mathcal V_m$ that meet at 
$P_0$ are in bijection with the set of roots of $P_0$ of multiplicity 
$\geq m$. In particular $\alpha_0$ correspond to one of these 
irreducible components; let us denote it by $\mathcal V_{m,\alpha_0}$. 
The algebraic variety $\mathcal V_{m,\alpha_0}$ is a \emph{a fortioti} 
$K$-manifold. Moreover there exists a differentiable function $f$ which 
is defined on a neighborhood of $P_0$ in $\mathcal V_{m,\alpha_0}$ and 
follows the root $\alpha_0$, \emph{i.e.} $f(P_0) = \alpha_0$ and for all 
$P$ such that $f(P)$ is defined, $\alpha = f(P)$ is a root of 
multiplicity (at least) $m$ of $P$. The existence of $f$ follows from 
Hensel's Lemma applied to the factorisation $P_0(X) = (X-\alpha_0)^m 
B_0(X)$ where $B_0(X)$ is a polynomial which is coprime to $X - 
alpha_0$. We can finally compute the differential of $f$ at $P_0$ 
(along $\mathcal V_{m,\alpha_0}$) by differentiating the equality
$P^{(m-1})(\alpha) = 0$ where $P^{(i})$ denotes the $i$-th derivative
of $P$. We find:
\begin{equation}
\label{eq:dalphamult}
d \alpha = - \frac{dP^{(m-1)}(\alpha_0)}{P_0^{(m)}(\alpha_0)}.
\end{equation}

@@

>>>>

We now address the case of a multiple roots. Let $P_0$ be a monic 
polynomial of degree $d$ and $\alpha_0 \in K$ be a root of $P_0$ having 
multiplicity $m > 1$. Due to the multiplicity, it is no longer possible 
to follow to root $\alpha_0$ in a neighborhood of $P_0$. But it is if we 
restrict ourselves to polynomials which have a root of multiplicity $m$ 
close to $\alpha_0$. More precisely, let us consider the locus $\mathcal 
V_m$ consisting of monic polynomials of degree $d$ having a root of 
multiplicity at least $m$. It is a Zariski close subset of $K_d[X]$ and 
it contains by assumption the polynomial $P_0$ we started with. 
Moreover, the irreducible components of $\mathcal V_m$ that meet at 
$P_0$ are in bijection with the set of roots of $P_0$ of multiplicity 
$\geq m$. In particular $\alpha_0$ correspond to one of these 
irreducible components; let us denote it by $\mathcal V_{m,\alpha_0}$. 
The algebraic variety $\mathcal V_{m,\alpha_0}$ is a \emph{a fortioti} 
$K$-manifold. Moreover there exists a differentiable function $f$ which 
is defined on a neighborhood of $P_0$ in $\mathcal V_{m,\alpha_0}$ and 
follows the root $\alpha_0$, \emph{i.e.} $f(P_0) = \alpha_0$ and for all 
$P$ such that $f(P)$ is defined, $\alpha = f(P)$ is a root of 
multiplicity (at least) $m$ of $P$. The existence of $f$ follows from 
Hensel's Lemma applied to the factorisation $P_0(X) = (X-\alpha_0)^m 
B_0(X)$ where $B_0(X)$ is a polynomial which is coprime to $X - 
\alpha_0$. We can finally compute the differential of $f$ at $P_0$ 
(along $\mathcal V_{m,\alpha_0}$) by differentiating the equality
$P^{(m-1)}(\alpha) = 0$ where $P^{(i)}$ denotes the $i$-th derivative
of $P$. We find:
\begin{equation}
\label{eq:dalphamult}
d \alpha = - \frac{dP^{(m-1)}(\alpha_0)}{P_0^{(m)}(\alpha_0)}.
\end{equation}

@@

This phenomenon is general: if $P$ is a polynomial over $\Q_p$ known at 
some finite precision $O(p^N)$, a root $\alpha$ of $P$ having possibly 
multiplicity $m$ (\emph{i.e.} $P(\alpha), P'(\alpha), \ldots, P^{(m-1)} 
(\alpha)$ vanish at the given precision) can be computed at precision 
roughly $O(p^{N/m})$. However, if we know in advance that $\alpha$ has 
indeed multiplicity $m$ then we can compute it at precision $O(p^{N-c})$ 
where $c$ is some constant.
